The Shape Matching Element Method for Meshless Animation of NURBs Models

### 2021-01-23 paper/sections/methods.tex
performance table
multipart projection finally done, yeesh

\begin{figure}
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{Figure showing input model NURBS parts and shape matching to each part}
    \label{fig:shapematching}
\end{figure}

@@

@@

\begin{equation*}
    \begin{split}
        \tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \mathbbm{1}^T\wt_k \samp{\Jnurbs}_k\vc{q}_k \\
        \vc{c}_i^1 & = \left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)^{-1}\samp{\Pmat^1}^T\wt_i\left(\samp{\Jnurbs}_i\vc{q}_i-\mathbbm{1}\tr_i\right)\\
        \vc{c}_i^2 & = \left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)^{-1}\samp{\Pmat^2}^T\wt_i\left(\samp{\Jnurbs}_i\vc{q}_i-\samp{\Pmat^1}\vc{c}_i^1 - \mathbbm{1}\tr_i\right)
    \end{split},
    \label{eq:hfit} 
\end{equation*} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^1}$ is the stacked evaluation of this matrix at $n$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients. 
Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3n\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.

Coupling via the deformation origins ensures our method will reproduce rigid and linear deformations in regions around these points, while higher order 
terms help compensate for more local deformations.

@@

%%%% OLD STUFF BELOW HERE  %%%%%
In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface. Therefore for the entire model each shape has a boundary, a polynomial describing its deformation, as well as an associated center of mass. These centers of mass may not be distinct, meaning that a single center of mass may be shared among multiple surfaces. This sharing of center of masses is typical and in some cases required (\myworries{reference center of mass section that describes how for planar surfaces we need to share COM to get a COM out of plane}). This leads to a new definition of the center of mass where it is the mean of all the boundary points associated with it.

In defining the shape matching fitting problem, we will form a system similar to equation \ref{eqn:fitting_one}. In this case where we have $n$ shape elements our set of unknown coefficients is $\hat{\mathbf{c}} = \left[\mathbf{c}_1^T \dots \mathbf{c}_i^T \dots \mathbf{c}_n^T \right]^T \in \mathbb{R}^{3kn}$ where $\mathbf{c}_i$ corresponds to the unknown polynomial coefficients for the $i$-th shape element. The set of boundary positions is similarly written as $\hat{\mathbf{b}}=\left[\mathbf{b}_1^T \dots \mathbf{b}_i^T \dots \mathbf{b}_n^T \right]^T \in \mathbb{R}^{3N}$ where $N$ is the total number of boundary points across all elements. If we have $m$ centers of mass we write the set of polynomial constant terms as $\hat{\mathbf{t}} = \left[\mathbf{t}_1^T \dots \mathbf{t}_i^T \dots \mathbf{t}_m^T \right]^T \in \mathbb{R}^{3m}$. With these terms, we write the polynomial fitting system for the entire model as
\begin{equation}
\begin{pmatrix}
 \mathbf{A}_1 &      &			  &			&				&\mathbf{S}_1\\  
 &            \ddots &            &  		&				&\vdots\\ 
 &            &      \mathbf{A}_i &   		&				&\mathbf{S}_i\\
 &            &      &            \ddots	&				&\vdots\\
 &            &      &            &			\mathbf{A}_n	&\mathbf{S}_n
\end{pmatrix}
\begin{bmatrix}
 \hat{\mathbf{c}}  \\ 
 \hat{\mathbf{t}}   
\end{bmatrix}
=
\hat{\mathbf{b}}
\text{,}
\end{equation}
where $\mathbf{A}_i$ is the monomial basis matrix for the $i$-th shape element, and $\mathbf{S}_i$ is a selection matrix that extracts the $j$-th center of mass which is associated with the $i$-th shape. We simplify this as
\begin{equation}
\begin{pmatrix}
\hat{\mathbf{A}} & \hat{\mathbf{S}}
\end{pmatrix}
\begin{bmatrix}
 \hat{\mathbf{c}}  \\ 
 \hat{\mathbf{t}}   
\end{bmatrix}
=
\hat{\mathbf{b}}
\text{.}
\end{equation}
Like in the single element case, we set the $\hat{\mathbf{t}}$ to be equal to the mean of a set of boundary positions, but in this case each $\mathbf{t}_j$ is equal to the mean of only the boundary positions associated with the $j$-th center of mass. The definition of the values of $\hat{\mathbf{t}}$ is similar to \ref{eqn:t_solution}:
\begin{equation}
\hat{\mathbf{t}} =
\underbrace{
\begin{bmatrix}
\mathbf{T}_1 \\ 
\vdots       \\
\mathbf{T_m}
\end{bmatrix}}_{\hat{\mathbf{T}}}
\text{,}
\end{equation}

where the entries each $\mathbf{T}_i \in \mathbb{R}^{3N}$ take a similar form, but are only nonzero for columns corresponding to associated boundary vertices.

Finally, to solve for the full set of non-constant term coefficients, $\hat{\mathbf{c}}$, we solve in the same manner as equation \ref{eqn:c_one_element} where we have
\begin{equation}
\label{eqn:c_multiple_elements}
\begin{split}
\hat{\mathbf{c}} & = (\hat{\mathbf{A}}^T\hat{\mathbf{A}})^{-1}\hat{\mathbf{A}}^T(\mathbf{I-\hat{S}\hat{T})\hat{\mathbf{b}}} \\
           & = \mathbf{L}\hat{\mathbf{b}}
\end{split}
\text{.}
\end{equation}

\dave{this is good but we shouldn't imply we solve the system above since we do Hierarchical ordering}
\paragraph*{Hierarchical fitting}
A problem may emerge in the shape matching process where the system may be underdetermined. To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation. The produces a set of smaller systems which are guaranteed to be determined \myworries{probably need to justify this more}. The intuition behind this can be described by the \textit{hierarchical ordering principle}, which argues that the behavior of a system is typically dominated by lower order effects \myworries{cite something here.}. To describe our fitting strategy, we begin with the following shape matching system for a single element
\begin{equation}
\label{eqn:fitting_one_copy}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{.}
\end{equation}
We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$. Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively. First we construct a system for the linear terms:
\begin{equation}
\label{eqn:fitting_linear}
\left[
\mathbf{A}_{\text{lin}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{lin}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
which leads to the following expression for the linear coefficients
\begin{equation}
\label{eqn:c_linear}
\begin{split}
\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}_{\text{lin}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
\begin{equation}
\label{eqn:fitting_quadratic}
\left[
\mathbf{A}_{\text{quad}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{quad}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
\text{.}
\end{equation}
Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
\begin{equation}
\label{eqn:c_quad}
\begin{split}
\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
           & = \mathbf{L}_{\text{quad}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
The final form for the full $\mathbf{L}$ matrix is thus given by
\begin{equation}
\mathbf{L} =
\begin{bmatrix}
\mathbf{L}_{\text{lin}} \\
\mathbf{L}_{\text{quad}} \\
\mathbf{T}
\end{bmatrix}
\end{equation}

The recursive nature of this process permits a more general description for arbitrary orders, which we write in algorithm \myworries{TODO}.

\subsection{Blending Weights}
\dave{Desireable Properties}
\dave{Construction via post normalization}
\dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
\dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
\dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
All that remains for the deformation map is to compute blending weights. Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest. This idea lends itself to a simple distance-based weighting method. The first step in this process is to compute \textit{distance weights}. For a given point $\mathbf{X}$ and a set of $n$ we wish to compute a set of distance weights to each surface $\mathbf{\theta}(\mathbf{X}) = \left(\theta_1, \dots, \theta_i, \dots, \theta_n \right)$.

>>>>

@@

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.}
    \label{fig:multiparts}
\end{figure}

@@

\begin{equation}
    \begin{split}
        \tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \underbrace{\mathbbm{1}^T\wt_k \samp{\Jnurbs}_k}_{B_{ik}}\vc{q}_k \\
        \vc{c}_i^1 & = \underbrace{\left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)}_{A^1_i}^{-1}\bigg(\underbrace{\samp{\Pmat^1}^T\wt_i\samp{\Jnurbs}_i}_{B^1_i}\vc{q}_i-\underbrace{\samp{\Pmat^1}^T\wt_i\mathbbm{1}}_{T^1_i}\tr_i\bigg)\\
        \vc{c}_i^2 & = \underbrace{\left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)}_{A^2_i}^{-1}\bigg(\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Jnurbs}_i}_{B^2_i}\vc{q}_i-\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Pmat^1}}_{D^2_i}\vc{c}_i^1 - \underbrace{\samp{\Pmat^2}^T\wt_i\mathbbm{1}}_{T^2_i}\tr_i\bigg)
    \end{split},
    \label{eq:hfit} 
\end{equation} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^1}$ is the stacked evaluation of this matrix at $n$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients. 
Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3n\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.

%%single projector for every object
When assembled for all parts,~\refeq{hfit}, yields a block upper triangular matrix $U$ and a spare matrix $B$ ~(\reffig{multiparts}) which can be used to efficiently to compute the projection 
operator for the entire object $\Pi = U^{-1}B\vc{q}$. 
The structure of $U$ and $B$ ensures that $\Pi$ only couples objects which share deformation centers, which implies a sparse $\Pi$. 
The per-part projection operators $\Pi_i$ correspond to blocks of rows of $\Pi$. 
Coupling via the deformation origins ensures our method will reproduce rigid and linear deformations in regions around these points, while higher order 
terms help compensate for more local deformations.

@@

%%%% OLD STUFF BELOW HERE  %%%%%
\subsection{Blending Weights}
\dave{Desireable Properties}
\dave{Construction via post normalization}
\dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
\dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
\dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
All that remains for the deformation map is to compute blending weights. Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest. This idea lends itself to a simple distance-based weighting method. The first step in this process is to compute \textit{distance weights}. For a given point $\mathbf{X}$ and a set of $n$ we wish to compute a set of distance weights to each surface $\mathbf{\theta}(\mathbf{X}) = \left(\theta_1, \dots, \theta_i, \dots, \theta_n \right)$.
